---
title: "Predictors report"
author: "Carles Milà"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center', echo=F)

library("targets")
library("tidyverse")
library("stars")
library("viridis")
library("RColorBrewer")
library("sf")
tar_load("cat")
cat <- read_rds(cat)
```

# Spatial predictors

## Coordinates

```{r coords, fig.height=3, fig.width=7}
tar_load("coords")
coords <- read_rds(coords)
par(mfrow=c(1,2))
plot(coords[1,,], breaks = "equal", col=viridis(8), main="x", 
     key.pos = NULL, reset = FALSE, add.geom=cat)
plot(coords[2,,], breaks = "equal", col=viridis(8), main="y", 
     key.pos = NULL, reset = FALSE, add.geom=cat)
par(mfrow=c(1,1))
rm("coords")
```

## Elevation

```{r dem, fig.height=3, fig.width=5}
tar_load("terrain")
terrain <- read_rds(terrain)
plot(terrain[1], breaks = "equal", col = terrain.colors(10), main="", add.geom=cat)
```

## Slope

```{r slope, fig.height=3, fig.width=5}
plot(terrain[2], breaks = "equal", col = terrain.colors(10), main="", add.geom=cat)
```

## Aspect

```{r aspect, fig.height=3, fig.width=5}
plot(terrain[3], breaks = "equal", col = terrain.colors(10), main="", add.geom=cat)
rm("terrain")
```

## Impervious density (IMD)

```{r imd, fig.height=3, fig.width=5}
tar_load("imd")
imd <- read_rds(imd)
plot(imd, breaks ="equal", main="", col = heat.colors(10, rev=T), add.geom=cat)
rm("imd")
```

## Tree Cover Density (TCD)

```{r tcd, fig.height=3, fig.width=5}
tar_load("tcd")
tcd <- read_rds(tcd)
plot(tcd, breaks ="equal", main="", col=brewer.pal(8, "Greens"), add.geom=cat)
rm("tcd")
```

## Distance to coast

```{r coast, fig.height=3, fig.width=5}
tar_load("coast")
coast <- read_rds(coast)
plot(coast, breaks ="equal", main="", col=rev(brewer.pal(8, "YlGnBu")), 
     add.geom=cat)
rm("coast")
```

## Land use - Land cover

```{r lulc, fig.height=3, fig.width=5}
tar_load("lulc")
lulc <- read_rds(lulc)
plot(lulc["agriculture",,], breaks ="equal", main="", 
     col=brewer.pal(5, "Oranges"), add.geom=cat)
rm("lulc")
```

## Industrial point sources

```{r psources`, fig.height=3, fig.width=5}
tar_load("psources")
psources <- read_rds(psources)
plot(psources[1,,], breaks ="equal", main="", 
     col=brewer.pal(8, "Oranges"), add.geom=cat)
rm("psources")
```

## Population denisty

```{r popu, fig.height=3, fig.width=5}
tar_load("popu")
popu <- read_rds(popu)
plot(popu, breaks = c(0, 1, 10, 50, 100, 4427), main="",
     col=mako(5), add.geom=cat)
rm("popu")
```

## Road density

```{r roaddens, fig.height=3, fig.width=5}
tar_load("roaddens")
roaddens <- read_rds(roaddens)
plot(roaddens[3,,], col=magma(7), breaks = c(0,1,10,50,100,200,500,1200),
     main="", add.geom=cat)
rm("roaddens")
```

# Temporal predictors

# Spatio-temporal predictors

## NDVI

```{r ndvi, fig.height=3, fig.width=9}
tar_load("ndvi")
ndvi <- read_rds(ndvi$ndvi18)
plot(ndvi[,,,1:4], col=brewer.pal(8, "Greens"), main="", 
     mfrow=c(1,4), add.geom=cat)
rm("ndvi")
```

## NTLI

```{r ntli, fig.height=3, fig.width=9}
tar_load("ntli")
ntli <- read_rds(ntli$ntli18)
plot(ntli, col=viridis(8), main="", mfrow=c(1,3), add.geom=cat)
rm("ntli")
```

# Reanalysis predictors

## ERA5

```{r era5 rean, fig.width=9, fig.height=2}
tar_load("era5rean")
era5rean <- read_rds(era5rean$era5rean18)
par(mfrow=c(1,2))
plot(era5rean[1,,,1], col=viridis(8), main="Mean sea level pressure", 
     key.pos = NULL, reset = FALSE, add.geom=cat)
plot(era5rean[2,,,1], col=viridis(8), main="Planetary boundary height", 
     key.pos = NULL, reset = FALSE, add.geom=cat)
par(mfrow=c(1,1))
rm("era5rean")
```

## ERA5Land

```{r era5land rean, fig.width=9, fig.height=2}
tar_load("era5land")
era5land <- read_rds(era5land$era5land18)
par(mfrow=c(1,3))
plot(era5land[2,,,1], col=viridis(8), main="wind u component", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,3))
plot(era5land[3,,,1], col=viridis(8), main="wind v component", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,3))
plot(era5land[4,,,1], col=viridis(5), main="Total precipitation", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,3))
par(mfrow=c(1,1))
par(mfrow=c(1,4))
plot(era5land[1,,,1], col=viridis(8), main="average temperature", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,3))
plot(era5land[5,,,1], col=viridis(8), main="minimum temperature", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,3))
plot(era5land[6,,,1], col=viridis(8), main="maximum temperature", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,3))
plot(era5land[10,,,1], col=viridis(8), main="skin temperature (10h)", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,3))
par(mfrow=c(1,1))
rm("era5land")
```

## CAMSglobal

```{r CAMSglobal, fig.width=8, fig.height=3}
tar_load("camsglobal")
camsglobal <- read_rds(camsglobal$camsglobal18)
par(mfrow=c(1,4))
plot(camsglobal[2,,,1], col=viridis(8), main="AOD550 9h", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,4))
plot(camsglobal[1,,,1], col=viridis(8), main="AOD550 12h", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,4))
plot(camsglobal[3,,,1], col=viridis(8), main="NO2 12h", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,4))
plot(camsglobal[4,,,1], col=viridis(5), main="O3 12h", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,4))
par(mfrow=c(1,1))
rm("camsglobal")
```

## CAMSrean

```{r CAMSrean, fig.width=8, fig.height=3}
tar_load("camsrean")
camsrean <- read_rds(camsrean$camsrean18)
par(mfrow=c(1,4))
plot(camsrean[1,,,1], col=viridis(8), main="PM25", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,4))
plot(camsrean[2,,,1], col=viridis(8), main="PM10", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,4))
plot(camsrean[3,,,1], col=viridis(8), main="NO2", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,4))
plot(camsrean[4,,,1], col=viridis(5), main="O3", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,4))
par(mfrow=c(1,1))
rm("camsrean")
```

## CAMSanaly

```{r CAMSanaly, fig.width=8, fig.height=3}
tar_load("camsanaly")
camsanaly <- read_rds(camsanaly$camsanaly19)
par(mfrow=c(1,4))
plot(camsanaly[1,,,1], col=viridis(8), main="NO2", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,4))
plot(camsanaly[2,,,1], col=viridis(8), main="PM10", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,4))
plot(camsanaly[3,,,1], col=viridis(8), main="PM25", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,4))
plot(camsanaly[4,,,1], col=viridis(5), main="O3", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,4))
par(mfrow=c(1,1))
rm("camsanaly")
```

# Atmospheric RS predictors

## MAIAC

```{r maiac, fig.height=3, fig.width=4}
tar_load("maiac")
maiac <- read_rds(maiac$maiac18)
plot(maiac[1,,,176], col=cividis(4), key.pos = 4, add.geom=cat,
     breaks=c(0,0.05,0.1,0.2,0.4)) 
plot(maiac[2,,,176], col=cividis(4), key.pos = 4, add.geom=cat,
     breaks=c(0,0.05,0.1,0.2,0.4))
rm("maiac")
```

## TROPOMI Tropospheric NO2

```{r tropomi no2, fig.height=3, fig.width=4}
tar_load("tropomi_no2")
tropomi_no2 <- read_rds(tropomi_no2$tropomi_no2_18)
plot(tropomi_no2[,,,58], col=cividis(5), key.pos = 4, add.geom=cat,
     breaks=c(-6.240e+14, 6.277e+14, 1.080e+15, 1.551e+15, 2.5e+15, 5.74e+15)) 
rm("tropomi_no2")
```

## OMI Tropospheric NO2

```{r omi no2, fig.height=3, fig.width=4}
tar_load("omi_no2")
omi_no2 <- read_rds(omi_no2$omi_no2_18)
plot(omi_no2[,,,176], col=cividis(5), key.pos = 4, add.geom=cat,
     breaks=c(-6.240e+14, 6.277e+14, 1.080e+15, 1.551e+15, 2.5e+15, 5.74e+15))
rm("omi_no2")
```

## TROPOMI Total O3

```{r tropomi o3, fig.height=3, fig.width=4}
tar_load("tropomi_o3")
tropomi_o3 <- read_rds(tropomi_o3$tropomi_o3_18)
plot(tropomi_o3[,,,58], col=cividis(6), key.pos = 4, add.geom=cat,
     breaks=c(300, 305, 310, 315, 320, 325, 330))
rm("tropomi_o3")
```

## OMI Total O3

```{r omi o3, fig.height=3, fig.width=4}
tar_load("omi_o3")
omi_o3 <- read_rds(omi_o3$omi_o3_18)
plot(omi_o3[,,,176], col=cividis(6), key.pos = 4, add.geom=cat,
     breaks=c(300, 305, 310, 315, 320, 325, 330))
rm("omi_o3")
```

## MODIS LST

```{r modis lst, fig.width=9, fig.height=3}
tar_load("lst_terra")
lst_terra <- read_rds(lst_terra$lst18)
tar_load("lst_aqua")
lst_aqua <- read_rds(lst_aqua$lst18)
par(mfrow=c(1,4))
plot(lst_terra[1,,,1], col=viridis(8), main="Terra day", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,4))
plot(lst_terra[2,,,1], col=viridis(8), main="Terra night", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,4))
plot(lst_aqua[1,,,1], col=viridis(8), main="Aqua day", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,4))
plot(lst_aqua[2,,,1], col=viridis(5), main="Aqua night", 
     key.pos = NULL, reset = FALSE, add.geom=cat, mfrow=c(1,4))
par(mfrow=c(1,1))
rm("lst_terra", "lst_aqua")
```

