---
title: "Station exploratory analysis"
author: "Carles Milà"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.align = 'center')

library("tidyverse")
library("sf")
library("knitr")
library("targets")
library("lubridate")
library("ggthemes")
library("lme4")
library("broomExtra")
library("cowplot")
library("plotly")
```

# Air pollution

```{r ap measures}
# Read station and boundary data
tar_load("ap_stations")
readings <- bind_rows(read_rds(ap_stations$readings18),
                      read_rds(ap_stations$readings19),
                      read_rds(ap_stations$readings20))
stations <- read_rds(ap_stations$stations)
tar_load("cat")
cat <- read_rds(cat)
```

## Number of stations and measurements

```{r measurement availability, fig.width=9, fig.height=3}
readings_tab <- readings %>%
  group_by(pollutant, ID) %>%
  summarise(N=n(), .groups = "keep") %>%
  ungroup()
ggplot(readings_tab) +
  geom_histogram(aes(N), colour = "grey50") +
  facet_wrap(~ pollutant, nrow = 1) +
  theme_bw()
readings_tab <-  readings_tab %>%
  group_by(pollutant) %>%
  summarise(`Mean number of days (SD)` = paste0(round(mean(N, 1)),
                                                " (", round(sd(N, 1)), ")"),
            `Number of stations` = n())
kable(readings_tab)
```

## Station type

```{r station types, fig.width=7, fig.height=6}
# Get unique tuples ID-pollutant
IDs_pollutant <- dplyr::select(readings, ID, pollutant, measure_type)
IDs_pollutant <- IDs_pollutant[!duplicated(IDs_pollutant),]
IDs_pollutant <- left_join(IDs_pollutant, stations, by = "ID")

# Table
IDs_pollutant_tab <- IDs_pollutant %>%
  dplyr::select(ID, pollutant, st_type) %>%
  group_by(pollutant, st_type) %>%
  summarise(N1 = n(), .groups="keep") %>%
  group_by(pollutant) %>%
  mutate(N2=sum(N1),
         `N (%)` = paste0(N1, " (", round(N1/N2*100, 1), "%)")) %>%
  ungroup() %>%
  dplyr::select(-N1, -N2) %>%
  pivot_wider(names_from = "st_type", values_from = "N (%)")
kable(IDs_pollutant_tab)

# Map
ggplot() +
  geom_sf(data=cat, alpha=0.2) +
  geom_sf(data=IDs_pollutant, aes(colour=st_type, geometry=geometry), size=1) +
  facet_wrap(~ pollutant, nrow=2) +
  theme_bw() + 
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  scale_colour_colorblind() + labs(colour="Station type")
```

## Station measure type

```{r station measure types, fig.width=7, fig.height=6}
# Table
IDs_pollutant_tab <- IDs_pollutant %>%
  dplyr::select(ID, pollutant, measure_type) %>%
  group_by(pollutant, measure_type) %>%
  summarise(N1 = n(), .groups="keep") %>%
  group_by(pollutant) %>%
  mutate(N2=sum(N1),
         `N (%)` = paste0(N1, " (", round(N1/N2*100, 1), "%)")) %>%
  ungroup() %>%
  dplyr::select(-N1, -N2) %>%
  pivot_wider(names_from = "measure_type", values_from = "N (%)")
kable(IDs_pollutant_tab)

# Map
ggplot() +
  geom_sf(data=cat, alpha=0.2) +
  geom_sf(data=IDs_pollutant, aes(colour=measure_type, geometry=geometry), size=1) +
  facet_wrap(~ pollutant, nrow=2) +
  theme_bw() + 
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  scale_colour_colorblind() + labs(colour="Station type")
```

## Yearly averages

```{r yearly conc}
yearly_avgs <- readings %>%
  mutate(year = year(as.Date(date))) %>%
  group_by(year, pollutant, ID) %>%
  summarise(conc=mean(conc), .groups="keep") %>%
  group_by(year, pollutant) %>%
  summarise(conc=mean(conc), .groups="keep") %>%
  pivot_wider(names_from = "year", values_from = "conc")
kable(yearly_avgs, digits=1)
```

```{r yearly maps, fig.height=6, fig.width=10}
yearly_avgs <- readings %>%
  mutate(year = year(as.Date(date))) %>%
  group_by(year, pollutant, ID) %>%
  summarise(conc=mean(conc), .groups="keep") 
yearly_avgs <- left_join(yearly_avgs, stations, by = "ID")

p1 <- ggplot() +
  geom_sf(data=cat, alpha=0.2) +
  geom_sf(data=dplyr::filter(yearly_avgs, pollutant=="NO2"),
                             aes(colour=conc, geometry=geometry), size=1) +
  facet_wrap(~ year) +
  theme_bw() + 
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        legend.position = "bottom") +
  labs(colour="Average concentration") + 
  scale_colour_viridis_c() +  ggtitle("NO2")

p2 <- ggplot() +
  geom_sf(data=cat, alpha=0.2) +
  geom_sf(data=dplyr::filter(yearly_avgs, pollutant=="O3"),
                             aes(colour=conc, geometry=geometry), size=1) +
  facet_wrap(~ year) +
  theme_bw() + 
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        legend.position = "bottom") +
  labs(colour="Average concentration") + 
  scale_colour_viridis_c() +  ggtitle("O3")

p3 <- ggplot() +
  geom_sf(data=cat, alpha=0.2) +
  geom_sf(data=dplyr::filter(yearly_avgs, pollutant=="PM2.5"),
                             aes(colour=conc, geometry=geometry), size=1) +
  facet_wrap(~ year) +
  theme_bw() + 
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        legend.position = "bottom") +
  labs(colour="Average concentration") + 
  scale_colour_viridis_c() +  ggtitle("PM2.5")

p4 <- ggplot() +
  geom_sf(data=cat, alpha=0.2) +
  geom_sf(data=dplyr::filter(yearly_avgs, pollutant=="PM10"),
                             aes(colour=conc, geometry=geometry), size=1) +
  facet_wrap(~ year) +
  theme_bw() + 
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        legend.position = "bottom") +
  labs(colour="Average concentration") + 
  scale_colour_viridis_c() +  ggtitle("PM10")

plot_grid(p1, p2, p3, p4, nrow=2)
```

## Temporal and spatial variance components 

We fit a LMM to the log(exposure) with a day and station indicator random intercept and
we extract the temporal and spatial ICC (%) by pollutant.

```{r icc}
readings_log <- mutate(readings, conc=log(conc))

icc <- function(df){
  mod <-lmer(conc ~ (1|ID) + (1|date), data = df)
  res <- tidy(mod) %>%
    dplyr::filter(effect=="ran_pars") %>%
    mutate(s2 = estimate^2) 
  res$ICC <- res$s2/sum(res$s2)*100
  dplyr::select(res, group, ICC)
}

# Compute
no2_icc <- icc(dplyr::filter(readings, pollutant=="NO2"))
o3_icc <- icc(dplyr::filter(readings, pollutant=="O3"))
pm25_icc <- icc(dplyr::filter(readings, pollutant=="PM2.5"))
pm10_icc <- icc(dplyr::filter(readings, pollutant=="PM10"))

# Merge and table
res_icc <- bind_rows(mutate(no2_icc, pollutant="NO2"), 
                     mutate(o3_icc, pollutant="O3"), 
                     mutate(pm25_icc, pollutant="PM2.5"), 
                     mutate(pm10_icc, pollutant="PM10")) %>%
  mutate(group = ifelse(group=="date", "temporal", ifelse(group=="ID", "spatial", "residual"))) %>%
  pivot_wider(names_from = "group", values_from = "ICC")

kable(res_icc, digits=1)
```

## Time series

Interactive plot: Zoom in and pan to see the detail. Each line is a different
station.

```{r ts, fig.width=10, fig.height=3}
readings <- mutate(readings, 
                   date = as.Date(date),
                   ID = as.character(ID))
no2 <- dplyr::filter(readings, pollutant == "NO2") %>%
  complete(date = seq.Date(min(date), max(date), by="day"),
           nesting(ID))

# NO2
no2 <- dplyr::filter(readings, pollutant == "NO2") %>%
  complete(date = seq.Date(min(date), max(date), by="day"),
           nesting(ID))
p1 <- ggplot(no2) + 
  geom_line(aes(x=date, y=conc, colour=ID)) +
  ylab("Concentration") +
  theme_bw() + theme(legend.position = "none") + ggtitle("NO2")
ggplotly(p1)

# O3
o3 <- dplyr::filter(readings, pollutant == "O3") %>%
  complete(date = seq.Date(min(date), max(date), by="day"),
           nesting(ID))
p2 <- ggplot(o3) + 
  geom_line(aes(x=date, y=conc, colour=ID)) +
  ylab("Concentration") +
  theme_bw() + theme(legend.position = "none") + ggtitle("O3")
ggplotly(p2)

# PM10
pm10 <- dplyr::filter(readings, pollutant == "PM10") %>%
  complete(date = seq.Date(min(date), max(date), by="day"),
           nesting(ID))
p3 <- ggplot(pm10) + 
  geom_line(aes(x=date, y=conc, colour=ID)) +
  ylab("Concentration") +
  theme_bw() + theme(legend.position = "none") + ggtitle("PM10")
ggplotly(p3)

# PM2.5
pm25 <- dplyr::filter(readings, pollutant == "PM2.5") %>%
  complete(date = seq.Date(min(date), max(date), by="day"),
           nesting(ID))
p4 <- ggplot(pm25) + 
  geom_line(aes(x=date, y=conc, colour=ID)) +
  ylab("Concentration") +
  theme_bw() + theme(legend.position = "none") + ggtitle("PM2.5")
ggplotly(p4)
```

# Temperature

```{r meteo measures}
# Read station and boundary data
tar_load("meteo_stations")
readings <- bind_rows(read_rds(meteo_stations$readings18),
                      read_rds(meteo_stations$readings19),
                      read_rds(meteo_stations$readings20))
stations <- read_rds(meteo_stations$stations)
```

## Number of stations and measurements

```{r measurement availability temp}
readings_tab <- readings %>%
  group_by(ID, entity) %>%
  summarise(N=n(), .groups = "keep") %>%
  ungroup() %>%
  group_by(entity) %>%
  summarise(`Mean number of days (SD)` = paste0(round(mean(N, 1)), 
                                                " (", round(sd(N, 1)), ")"),
            `Number of stations` = n())
kable(readings_tab)
```

```{r station types temp, fig.width=4, fig.height=3}
ggplot() +
  geom_sf(data=cat, alpha=0.2) +
  geom_sf(data=stations, aes(colour=entity, geometry=geometry), size=1) +
  theme_bw() + 
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  scale_colour_colorblind() 
```

## Yearly averages

```{r yearly avg temp}
yearly_avgs <- readings %>%
  mutate(year = year(as.Date(date))) %>%
  group_by(year, ID) %>%
  summarise(tavg=mean(tavg, na.rm=T),
            tmax=mean(tmax, na.rm=T), 
            tmin=mean(tmin, na.rm=T), .groups="keep") %>%
  group_by(year) %>%
  summarise(tavg=mean(tavg, na.rm=T), 
            tmax=mean(tmax, na.rm=T), 
            tmin=mean(tmin, na.rm=T), .groups="keep")
kable(yearly_avgs, digits=1)
```

```{r yearly maps temp, fig.height=8, fig.width=9}
yearly_avgs <- readings %>%
  mutate(year = year(as.Date(date))) %>%
  group_by(year, ID) %>%
  summarise(tavg=mean(tavg, na.rm=T),
            tmax=mean(tmax, na.rm=T), 
            tmin=mean(tmin, na.rm=T), .groups="keep") 
yearly_avgs <- left_join(yearly_avgs, stations, by = "ID") %>%
  pivot_longer(cols=c("tavg", "tmax", "tmin"), names_to = "variable", values_to = "temp")

ggplot() +
  geom_sf(data=cat, alpha=0.2) +
  geom_sf(data=yearly_avgs, aes(colour=temp, geometry=geometry), size=1) +
  facet_grid(variable ~ year) +
  theme_bw() + 
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  labs(colour="ºC") + 
  scale_colour_distiller(palette = "Spectral")
```

## Temporal and spatial variance components 

We fit a LMM to the exposure with a day and station indicator random intercept and
we extract the temporal and spatial ICC (%) by pollutant.

```{r icc temp}
icc <- function(df, out){
  mod <-lmer(paste0(out, " ~ (1|ID) + (1|date)"), data = df)
  res <- tidy(mod) %>%
    dplyr::filter(effect=="ran_pars") %>%
    mutate(s2 = estimate^2) 
  res$ICC <- res$s2/sum(res$s2)*100
  dplyr::select(res, group, ICC)
}

# Compute
avg_icc <- icc(readings, "tavg")
min_icc <- icc(readings, "tmin")
max_icc <- icc(readings, "tmax")

# Merge and table
res_icc <- bind_rows(mutate(avg_icc, temp="avg"), 
                     mutate(min_icc, temp="min"), 
                     mutate(max_icc, temp="max")) %>%
  mutate(group = ifelse(group=="date", "temporal", ifelse(group=="ID", "spatial", "residual"))) %>%
  pivot_wider(names_from = "group", values_from = "ICC")

kable(res_icc, digits=1)
```

## Time series

Interactive plot: Zoom in and pan to see the detail. Each line is a different
station.

```{r ts temp, fig.width=10, fig.height=3}
# tavg
p1 <- ggplot(readings) +
  geom_line(aes(x=date, y=tavg, colour=ID)) +
  ylab("Average temp") +
  theme_bw() + theme(legend.position = "none")
ggplotly(p1)

# # tmin
# p1 <- ggplot(readings) + 
#   geom_line(aes(x=date, y=tmin, colour=ID)) +
#   ylab("Minimum temp") +
#   theme_bw() + theme(legend.position = "none")
# ggplotly(p1)
# 
# # tmax
# p1 <- ggplot(readings) + 
#   geom_line(aes(x=date, y=tmax, colour=ID)) +
#   ylab("Maximum temp") +
#   theme_bw() + theme(legend.position = "none")
# ggplotly(p1)
```
